---
title: "Finalizing"
author: "Kobe Sarausad"
date: "5/21/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{css, echo=FALSE}
@import url('https://fonts.googleapis.com/css2?family=Mukta:wght@200;400&display=swap');

.tbl {
  font-family: 'Mukta', sans-serif;
  width: 80%;
}

.tbl img {
  width: 30%;
  height: 30%;
}
```

```{r}
library(tidyverse)
library(kableExtra)
library(reactable)
library(reactablefmtr)
library(tidymodels)
library(rstantools)
library(rstanarm)
library(tidyverse)
library(vip)
library(glmnet)

set.seed(2022)

mvp <- read.csv("../data/cleanedmvp.csv")

mvp2022 <- read.csv("../data/cand2022.csv") %>% 
  select(-c(winshares, winsharesp48)) %>% 
  mutate(season = season + 1)

mvp2022 <- mvp2022 %>% 
  relocate(season, .after = rank)

mvp_full <- rbind(mvp, mvp2022)

mvp_full <- mvp_full %>% 
  relocate(player_id, .after = player)
```

### Introduction

Before we get started with any of this, we should acknowledge the players that receive hate from social media... This project was pretty fun in as I was able to quantify this hate by using text sentiment analysis to measure how much negativity comes up when people mention a specific player.

```{r}
mvp_full %>% 
  group_by(player, player_id) %>% 
  summarize(comp = sum(comp)) %>% 
  arrange(comp)
```



### Data Dictionary

The following are the variables that are in consideration for the model.

  - `rank` - ranking based on the number of votes the player received for that particular year
  - `season` - NBA season
  - `player` - NBA player
  - `player_id` - NBA player id (basketball reference)
  - `age` - NBA player's age for that particular season
  - `team` - NBA player's team for that particular season
  - `fpv` - first place votes
  - `mvppts` - points based on votes
  - `ptsmax` - total amount of points given to the pool of MVPs
  - `share` - the share of the max points the player received
  - `games` - number of games player played during the season
  - `mpg` - minutes played per game
  - `ppg` - points per game
  - `rbpg` - rebounds per game
  - `astpg` - assists per game
  - `stlpg` - steals per game
  - `blkpg` - blocks per game
  - `fgp` - field goal percentage
  - `threepp` - three point percentage
  - `ftp` - free throw percentage
  - `poss` - possessions played
  - `tmp` - total minutes played
  - `rap_o` - Points above average per 100 possessions added by player on offense, using both box and on-off components (FiveThirtyEight)
  - `rap_d` -	Points above average per 100 possessions added by player on defense, using both box and on-off components (FiveThirtyEight)
  - `rap_tot` - Points above average per 100 possessions added by player on both offense and defense, using both box and on-off components (FiveThirtyEight)
  - `war_tot` - Wins Above Replacement between regular season and playoffs (FiveThirtyEight)
  - `war_reg` - Wins Above Replacement for regular season (FiveThirtyEight)
  - `war_play` - Wins above Replacement for playoff season (FiveThirtyEight)
  - `pred_o` - Predictive points above average per 100 possessions added by player on offense (FiveThirtyEight)
  - `pred_d` - Predictive points above average per 100 possessions added by player on defense (FiveThirtyEight)
  - `pred_tot` - Predictive points above average per 100 possessions added by player on offense and defense (FiveThirtyEight)
  - `pace_impact` - Player impact on team possessions per 48 minutes (FiveThirtyEight)
  - `nposts` - number of relevant posts scraped on reddit
  - `avg_upvotes` - average upvotes per post
  - `avg_comments` - average number of comments per post
  - `comp` - composite sentiment score
  - `pos` - positive sentiment score
  - `neu` - neutral sentiment score
  - `neg` - negative sentiment score

## Modeling

### Feature Selection
  
We ran LASSO regression on the data (previous week) to determine which variables would be of good use to the model. As a result, LASSO regression ended up choosing the following as important variables: `age`, `ppg`, `rbpg`, `astpg`, `stlpg`, `fgp`, `tmp`, `war_tot`, `war_reg`, `pace_impact`, `pos`.

This gives us a good idea of what kinds of variables to use, in other words, it just points us in the right direction. So, using this information, we will remove some features whilst keeping others.

We'll first remove the following:
  - `war_play` since this relates to the impact on the court during the post-season, as we are trying to predict during the season
  - `war_tot` as well since it includes the playoff WAR
  - `nposts` since it should be around 25 for all players
  
Let's just begin with that.

### Spliting and Normalizing the data

```{r}
`%!in%` <- Negate(`%in%`)

features <- names(mvp_full)[-c(1, 2, 4, 5, 7:11, 22:23, 27, 29, 34)]

predict <- c(2, 7:11)

mvp_full[features[-1]] <- as.data.frame(scale(mvp_full[features[-1]]))

year_split <- sample(2010:2022, 9, FALSE)

mvp_train <- mvp_full %>%
  filter(season %in% year_split)

mvp_test <- mvp_full %>%
  filter(season %!in% year_split)
```

### Specify model

```{r}
boost_spec <- boost_tree(trees = 5000, tree_depth = 4) %>%
  set_engine("xgboost") %>%
  set_mode("regression")

boost_fit <- fit(boost_spec, mvppts ~ ., data = mvp_train[c(features, "mvppts")])
```

### Predicting on testing data

```{r}
boost_predict <- augment(boost_fit, new_data = mvp_test[c(features, "mvppts")])

boost_df <- cbind(boost_predict, mvp_test[c(2, 4)])

boost_df <- boost_df %>% 
  group_by(season) %>% 
  arrange(-.pred, .by_group = TRUE) %>% 
  mutate(pred_rank = 1:n(), rank = as.numeric(str_remove(rank, "T")), rank_diff_sq = (rank-pred_rank)^2, accuracy = pred_rank == rank, pred_share = .pred / sum(.pred))

boost_df %>% 
  group_by(season) %>% 
  summarize(mean(accuracy))
```

### Evaluating Error on share/pts vs. rank

```{r}

```

### Predicting on full dataset

```{r}
full_predict <- augment(boost_fit, new_data = mvp_full[c(features, "mvppts")])

full_df <- cbind(full_predict, mvp_full[c(2, 4)])

full_df <- full_df %>% 
  group_by(season) %>% 
  arrange(-.pred, .by_group = TRUE) %>% 
  mutate(pred_rank = 1:n(), rank = as.numeric(str_remove(rank, "T")), rank_diff_sq = (rank-pred_rank)^2, accuracy = pred_rank == rank)

sqrt(sum(full_df$rank_diff_sq))
sd(full_df$rank_diff_sq)

full_df %>% 
  group_by(season) %>% 
  summarize(mean(accuracy))
```

### Tabling the top 3 players predicted for every season (2010-2022)

```{r}
tab <- full_df %>% 
  group_by(season) %>% 
  arrange(-.pred, .by_group = TRUE) %>% 
  slice(1:3) %>% 
  mutate(pred_rank = 1:n(), rank = as.numeric(str_remove(rank, "T")), rank_diff_sq = (rank-pred_rank)^2, accuracy = pred_rank == rank) %>% 
  select(player, rank, pred_rank)

real_tab <- left_join(tab, mvp_full[c("season", "player", "player_id", "comp")]) %>% 
  mutate(img = paste0("https://www.basketball-reference.com/req/202106291/images/players/", player_id, ".jpg")) 

# real_tab %>% 
#   select(-c(player_id, prof_img)) %>% 
#   mutate(x = "") %>% 
#   relocate(x, .before = player) %>%
#   mutate(player = paste0(player, " (", season, ")")) %>% 
#   select(-1) %>% 
#   kable(align = "l", escape = FALSE, col.names = c("", "Player", "Actual Rank", "Predicted Rank", "Sent. Score")) %>% 
#   kable_material(full_width = F, c("striped")) %>% 
#   row_spec(0, color = "black") %>% 
#   column_spec(1, image = spec_image(real_tab$img, 60, 100), width = "4em") %>% 
#   column_spec(2, color = "black") %>% 
#   column_spec(3, color = "black", italic = T) %>% 
#   column_spec(4, color = "black", italic = T) %>% 
#   column_spec(5, background = spec_color(real_tab$comp, alpha = 1, begin = 0.3, end = 1,  option = "D", direction = 1), color = "white") 

table_react <- real_tab %>% 
  select(-player_id) %>% 
  relocate(img, .before = player) %>%
  mutate(player = ifelse(rank == 1, paste0(player, "\U1F3C6"), player), comp = round(comp, 2), season = ifelse(rank == pred_rank, paste0("\U2705", season), paste0("\U274C", season))) %>%
  rename("Predicted Rank" = "pred_rank") %>%
  rename("Sentiment" = "comp") %>% 
  reactable(
    class = "tbl",
    theme = fivethirtyeight(),
    columns = list(
      img = colDef(cell = embed_img()),
      Sentiment = colDef(style = pos_neg_colors("red", "green"))
      ),
    compact = TRUE,
    borderless = TRUE,
    striped = FALSE,
    fullWidth = FALSE,
    height = 490)
table_react
```

## Change of sampling method

```{r}
split_dat <- initial_split(mvp_full, 0.8)

mvp_train2 <- training(split_dat)
mvp_test2 <- testing(split_dat)

boost_fit2 <- fit(boost_spec, mvppts ~ ., data = mvp_train2[c(features, "mvppts")])

full_predict2 <- augment(boost_fit2, new_data = mvp_full[c(features, "mvppts")])

full_df2 <- cbind(full_predict2, mvp_full[c(2, 4)])

sqrt(sum(full_df2$rank_diff_sq))
sd(full_df2$rank_diff_sq)

full_df2 <- full_df2 %>% 
  group_by(season) %>% 
  arrange(-.pred, .by_group = TRUE) %>% 
  mutate(pred_rank = 1:n(), rank = as.numeric(str_remove(rank, "T")), rank_diff_sq = (rank-pred_rank)^2, accuracy = pred_rank == rank)

full_df2 %>% 
  group_by(season) %>% 
  summarize(mean(accuracy))

tab2 <- full_df2 %>% 
  group_by(season) %>% 
  arrange(-.pred, .by_group = TRUE) %>% 
  slice(1:3) %>% 
  mutate(pred_rank = 1:n(), rank = as.numeric(str_remove(rank, "T")), rank_diff_sq = (rank-pred_rank)^2, accuracy = pred_rank == rank) %>% 
  select(player, rank, pred_rank)

real_tab2 <- left_join(tab2, mvp_full[c("season", "player", "player_id", "comp")]) %>% 
  mutate(img = paste0("https://www.basketball-reference.com/req/202106291/images/players/", player_id, ".jpg")) 

table_react2 <- real_tab2 %>% 
  select(-player_id) %>% 
  relocate(img, .before = player) %>%
  mutate(player = ifelse(rank == 1, paste0(player, "\U1F3C6"), player), comp = round(comp, 2), season = ifelse(rank == pred_rank, paste0("\U2705", season), paste0("\U274C", season))) %>%
  rename("Predicted Rank" = "pred_rank") %>%
  rename("Sentiment" = "comp") %>% 
  reactable(
    class = "tbl",
    theme = fivethirtyeight(),
    columns = list(
      img = colDef(cell = embed_img()),
      Sentiment = colDef(style = pos_neg_colors("red", "green"))
      ),
    compact = TRUE,
    borderless = TRUE,
    striped = FALSE,
    fullWidth = FALSE,
    height = 490)
table_react2
```

```{r}
save.image(file = "final.RData")
```