---
title: 'MVP Prediction: Logistic Regression & Regularization (Week 2)'
author: "Kobe Sarausad"
date: "4/19/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
```

# Predicting the NBA MVP using season stats and Twitter sentiment

```{r}
library(tidyverse)
library(tidymodels)
library(nbastatR)
library(glmnet)
library(rtweet)
library(vader)
library(newsanchor)
library(httr)
library(rvest)

theme_cons <- theme_minimal() +
  theme(text = element_text(family = "Titillium Web"))

data <- read.csv("../data/fullmvp.csv")
raptor <- read.csv("https://raw.githubusercontent.com/fivethirtyeight/data/master/nba-raptor/historical_RAPTOR_by_player.csv")
```

## Begin by finding more season stats

```{r}
nba2021 <- seasons_schedule(2021)

test <- fanduel_summary(game_ids = nba2021$idGame[1])
players <- nba_players()

data$win <- data$Rank == 1
```

## EDA

```{r}
raptor %>% 
  ggplot() +
  geom_histogram(aes(x = raptor_total), bins = 100) +
  geom_vline(xintercept = mean(data$raptor_total), color = "orange") +
  geom_text(aes(x = mean(data$raptor_total) - 2, y = 3000, label = "MVP Candidates"), size = 3, family = "Titillium Web", angle = 90) +
  geom_vline(xintercept = mean(data$raptor_total[data$Rank == 1]), color = "seagreen4") +
    geom_text(aes(x = mean(data$raptor_total[data$Rank == 1]) + 2, y = 3000, label = "MVP"), size = 3, family = "Titillium Web", angle = -90) +
  labs(title = "MVPs on average have a higher raptor score") +
  theme_cons
```
```{r}
data %>% 
  ggplot() +
  geom_point(aes(x = Share, y = raptor_total, color = win)) +
  theme_cons +
  scale_color_manual(values = c("orange", "seagreen4"))
```

```{r}
data %>% 
  ggplot() +
  geom_point(aes(x = raptor_total, y = war_reg_season, color = Share)) +
  theme_cons +
  scale_color_gradient(low = "orange", high = "darkblue")  
```

```{r}
data %>% 
  ggplot() +
  geom_histogram(aes(x = Age), bins = 20) +
  theme_cons
```

```{r}
data %>% 
  group_by(Age) %>% 
  summarize(share = mean(Share)) %>% 
  ggplot() +
  geom_line(aes(x = Age, y = share)) +
  geom_point(data = data, aes(x = Age, y = Share), alpha = 0.2) +
  theme_cons
```



## Logistic Regression

Since we are using logistic regression, we need to make a binary variable which indicates if the player was the resulting MVP or not.

```{r}
split <- initial_split(data, strata = Share, prop = 0.5)
mvp_train <- training(split)
mvp_test <- testing(split)

glm.fit <- glm(win ~ Age + G + MP + PTS + TRB + AST + STL + BLK + FG. + X3P. + FT. + WS + WS.48 + raptor_offense + raptor_defense + raptor_total + war_reg_season + predator_offense + predator_defense + predator_total + pace_impact, data = mvp_train, family = 'binomial')
summary(glm.fit)
```

The p-values tell us some interesting things about how some features are important whilst some are not.

```{r}

```

## Ridge Regression

```{r}
features <- c("Age", "G", "MP", "PTS", "TRB", "AST", "STL", "BLK", "FG.", "X3P.", "FT.", "WS", "WS.48", "raptor_offense", "raptor_defense", "raptor_total", "war_reg_season", "predator_offense", "predator_defense", "predator_total", "pace_impact")

# lambda_seq <- 10^seq(2, -2, by = -.1)
# 
# fit <- glmnet(select(mvp_train, features), mvp_train$Share, alpha = 0, lambda  = lambda_seq)
# 
# ridge_cv <- cv.glmnet(select(mvp_train, features), mvp_train$Share, alpha = 0, lambda = lambdas_seq)
# best_lambda <- ridge_cv$lambda.min
# best_lambda
```

## LASSO Regression w/o standaridization

```{r}
mvp_train_ <- select(mvp_train, features, Share)

mvp_rec <- recipe(Share ~ ., data = mvp_train_) 

wf <- workflow() %>%
  add_recipe(mvp_rec)

mvp_boot <- bootstraps(mvp_train_)

tune_spec <- linear_reg(penalty = tune(), mixture = 1) %>%
  set_engine("glmnet")

lambda_grid <- grid_regular(penalty(), levels = 50)

set.seed(1234)
lasso_grid <- tune_grid(
  wf %>% add_model(tune_spec),
  resamples = mvp_boot,
  grid = lambda_grid
)

lasso_grid %>%
  collect_metrics()

lasso_grid %>%
  collect_metrics() %>%
  ggplot(aes(penalty, mean, color = .metric)) +
  geom_errorbar(aes(
    ymin = mean - std_err,
    ymax = mean + std_err
  ),
  alpha = 0.5
  ) +
  geom_line(size = 1.5) +
  facet_wrap(~.metric, scales = "free", nrow = 2) +
  scale_x_log10() +
  theme(legend.position = "none") + 
  cons_theme
```

## LASSO Regression w/ standaridization

```{r}
mvp_rec_std <- recipe(Share ~ ., data = mvp_train)  %>% 
  step_normalize(all_numeric(), -all_outcomes())

wf_std <- workflow() %>%
  add_recipe(mvp_rec_std)

mvp_boot_std <- bootstraps(mvp_train_)

tune_spec <- linear_reg(penalty = tune(), mixture = 1) %>%
  set_engine("glmnet")

lambda_grid <- grid_regular(penalty(), levels = 50)

set.seed(1234)
lasso_grid_std <- tune_grid(
  wf %>% add_model(tune_spec),
  resamples = mvp_boot_std,
  grid = lambda_grid
)

lasso_grid_std %>%
  collect_metrics()

lasso_grid_std %>%
  collect_metrics() %>%
  ggplot(aes(penalty, mean, color = .metric)) +
  geom_errorbar(aes(
    ymin = mean - std_err,
    ymax = mean + std_err
  ),
  alpha = 0.5
  ) +
  geom_line(size = 1.5) +
  facet_wrap(~.metric, scales = "free", nrow = 2) +
  scale_x_log10() +
  theme(legend.position = "none") +
  cons_theme
```
```{r}
lowest_rmse <- lasso_grid %>%
  select_best("rmse", maximize = FALSE)

final_lasso <- finalize_workflow(
  wf %>% add_model(tune_spec),
  lowest_rmse
)
```
## Regularization

```{r}

```

## Fetching some Lebron Tweets

```{r}
lebron <- search_tweets("Lebron", n = 1000, include_rts = FALSE, type = "popular")
```

##Sentiment Analysis

```{r}
get_vader("I love it")

get_vader("Mariners suck")

get_vader("Mariners suck but I still love them")

lebron$sentiment <- vader_df(lebron$text)

lebron$wsentscore <- lebron$sentiment$compound * lebron$favorite_count

lebron %>% 
  ggplot() +
  geom_point(aes(x = favorite_count, y = sentiment$compound)) +
  theme_cons

nytlebronmvp <- read_html("https://www.nytimes.com/2010/05/03/sports/basketball/03lebron.html") %>% 
  html_nodes(".css-g5piaz") %>% 
  html_text() %>% 
  paste(collapse = '')

get_vader(nytlebronmvp)
```